\subsection{Answers}

\begin{enumerate}
\item Which property or properties hold(s) for the track-before-detect method

\textbf{Answer:}

\begin{enumerate}
\item The approach can be used in tracking scenario with very low Signal-to-Noise-Ratio
\item Doesn't use a detector, input raw sensor data into MOT. 
\end{enumerate}
\item Which property or properties hold(s) for the point object tracking method?

\textbf{Answer:}

\begin{enumerate}
\item At most one detection per object per time scan.
\item Possibly provide object extent estimate
\end{enumerate}

\item Which property or properties hold(s) for the extended object tracking method?

\textbf{Answer:}

\begin{enumerate}
\item Possibly provide object extent estimate.
\item Possibly multiple measurements per object
\end{enumerate}

\item Which property or properties hold(s) for the group object tracking method?

\textbf{Answer:}

\begin{enumerate}
\item Possibly provide object extent estimate. correct .
\item Possibly multiple measurements per object
\item The tracked object consists of smaller "sub-objects" that move in common formation.
\end{enumerate}

\item Which property or properties hold(s) for the tracking with multi-path propagation method?

\textbf{Answer:}

\begin{enumerate}
\item Measurements resulted from multiple propagation paths.
\item Possibly multiple measurements per object.
\end{enumerate}

\item Which property or properties hold(s) for the tracking with unresolved measurements method?

\textbf{Answer:}

\begin{enumerate}
\item A group of close objects that collectively result in a single measurement
\end{enumerate}
\end{enumerate}

