\section{Answers to Questions}
\label{answer_questions}

{\textbf{Answer to questions in section \ref{introduction_self_driving_cars_questions}}}

\begin{enumerate}

\item Which of the following are components of longitudinal control? (Select all that apply)

\begin{itemize}
\item Braking and accelerating are components of longitudinal control. Thus, options 1 and 3 are correct
\end{itemize}

\item Which of the following is not an example of OEDR?

\begin{itemize}
\item Finding routes between locations is a long term planning problem and not OEDR. Hence, option 2 is correct.
\end{itemize}

\item Which of the following tasks would you expect a Level 2 system to perform? (Select all that apply)

\begin{itemize}
\item Options 1 and 3 are correct.
\end{itemize}

\item What is the distinction between Level 3 autonomy and Level 4 autonomy? 

\begin{itemize}
\item Option 4 is correct.  Level 3 systems cannot handle emergencies automatically and as a result require full user alertness.
\end{itemize}

\item What distinguishes Level 5 Autonomy from Level 4?

\begin{itemize}
\item Option 3 is correct. Level 5 systems can operate in any weather condition, on any road type or surface and in any scenario and remain safe.
\end{itemize}

\end{enumerate}

{\textbf{Answer to questions in section \ref{perception_questions}}}

\begin{enumerate}

\item Which of the following tasks are associated with perception? (Select all that apply) 
\begin{itemize}
\item Perception deals with position, motion estimation and object identification. Hence options 1 and 2 are correct.
\end{itemize}

\item Which of the following can be on road objects? (Select all that apply) 

\begin{itemize}
\item Potholes and vehicles are on road objects. So options 1 and 4 are correct.
\end{itemize}

\item Which of the following tasks pose challenges to perception? (Select all that apply) 

\begin{itemize}
\item All four choices pose challenges to perception
\end{itemize}

\item Which of the following sensors are used for ego localization? (Select all that apply) 


\begin{itemize}
\item Global Navigation Satellite System (GNSS) sensor provides position and velocity measurements, and can be used to estimate vehicle position and orientation for localization. An IMU provide acceleration and rotation rate measurements from accelerometers and gyroscopes, and can be used to estimate vehicle orientation and aid in localization in general. Thus, options 1 and four are correct.
\end{itemize}

\item Which of the following objects would be relevant for perception in adaptive cruise control?


\begin{itemize}
\item Adaptive cruise control detects vehicles ahead to control speed and to maintain safe driving distances. So
option 3 is correct. 
\end{itemize}

\end{enumerate}

