\section{Glossary}
\label{glossary}

\begin{itemize}
\item ACC Adaptive Cruise Control: A cruise control system for vehicles which controls longitudinal speed. 
ACC can maintain a desired reference speed or adjust its speed accordingly to maintain safe driving distances to other vehicles.
\end{itemize}

\begin{itemize}
\item FMEA Failure Mode and Effects Analysis: A bottom up approach of failure analysis which examines individual 
causes and determines their effects on the higher level system..
\end{itemize}

\begin{itemize}
\item GNSS Global Navigation Satellite System: A generic term for all satellite systems which provide position estimation. 
The Global Positioning System (GPS) made by the United States is a type of GNSS. 
Another example is the Russian made GLONASS (Globalnaya Navigazionnaya Sputnikovaya Sistema).
\end{itemize}

\begin{itemize}
\item HAZOP Hazard and Operability Study:A variation of FMEA (Failure Mode and Effects Analysis) which uses 
guide words to brainstorm over sets of possible failures which can arise..
\end{itemize}

\begin{itemize}
\item IMU Inertial Measurement Unit:A sensor device consisting of an accelerometer and a gyroscope. 
The IMU is used to measure vehicle acceleration and angular velocity, 
and its data can be fused with other sensors for state estimation..
\end{itemize}

\begin{itemize}
\item LIDAR Light Detection and Ranging:A type of sensor which detects range by transmitting 
light and measuring return time and shifts of the reflected signal.
\end{itemize}

\begin{itemize}
\item LTI Linear Time Invariant: A linear system whose dynamics do not change with time. 
For example, a car using the unicycle model is a LTI system. 
If the model includes the tires degrading over time (and changing the vehicle dynamics), then the system would no longer be LTI.
\end{itemize}

\begin{itemize}
\item LQR Linear Quadratic Regulation: A method of control utilizing full state feedback. The method seeks to optimize a quadratic cost function dependent on the state and control input.
\end{itemize}

\begin{itemize}
\item MPC Model Predictive Control: A method of control whose control input optimizes a user defined cost function over a finite time horizon. 
A common form of MPC is finite horizon LQR (linear quadratic regulation).
\end{itemize}

\begin{itemize}
\item ODD Operational Design Domain: The set of conditions under which a given system is designed to function. 
For example, a self driving car can have a control system designed for driving in urban environments, and another for driving on the highway..
\end{itemize}

\begin{itemize}
\item OEDR Object and Event Detection and Response: The ability to detect objects and events that 
immediately affect the driving task, and to react to them appropriately.
\end{itemize}


\begin{itemize}
\item PID Proportional Integral Derivative Control: A common method of control defined by 3 gains.
\begin{enumerate}
\item A proportional gain which scales the control output based on the amount of the error
\item An integral gain which scales the control output based on the amount of accumulated error
\item A derivative gain which scales the control output based on the error rate of change
\end{enumerate}
\end{itemize}


\begin{itemize}
\item RADAR Radio Detection And Ranging: A type of sensor which detects range and movement by transmitting 
radio waves and measuring return time and shifts of the reflected signal.
\end{itemize}

\begin{itemize}
\item SONAR Sound Navigation And Ranging: A type of sensor which detects range and movement by transmitting sound waves and measuring return time and shifts of the reflected signal. 
\end{itemize}