\section{Answers to Questions}
\label{answers_hardware_for_self_driving_cars}

{\textbf{Answer to questions in section \ref{questions_hardware_for_self_driving_cars}}}

\begin{enumerate}
\item What are the differences between exteroceptive sensors and proprioceptive sensors? (Select all that apply)
\item Which of the following exteroceptive sensors would you use in harsh sunlight?
\item Why is synchronization and timing accuracy important in the self driving system? Choose the primary reason.
\item Your autonomous vehicle is driving on the German autobahn at $150km/h$ and you wish to maintain safe following distances with other vehicles. 
		Assuming a safe following distance of $2s$,  what is the distance (in m) required between vehicles? Round your answer to 2 decimal places.

\item Using the same speed of $150km/h$, what is the braking distance (in m) required for emergency stops? 
	  Assume an aggressive deceleration of $5m/s^2$. Round your answer to two decimal places.
\item Suppose your vehicle was using long range cameras for sensing forward distance, 
	but it is now nighttime and the images captured are too dark. Which of the following sensors can be used to compensate?
\item What are the differences between an occupancy grid and a localization map? (Select all that apply)
\item The vehicle steps through the software architecture and arrives at the controller stage. 
		What information is required for the controller to output its commands to the vehicle?
\item What is (are) the role(s) of the system supervisor? (Select all that apply) 
\item Which of the following tasks should be assigned to the local planner?
\item What common objects in the environment appear in the occupancy grid?
\item Which of the following maps contain roadway speed limits?
\end{enumerate}

