\section{Camera Calibration}
\label{camera_calibration}

Section \ref{camera_projective_geometry_problem_definition} discussed which camera parameters are needed for projective
geometry to work. In this section, we will learn how to get these camera parameters using the mathematical tools
of calibration, see \cite{CameraCalibration1}. 

Thus, camera calibration is the process of estimating intrinsic and/or extrinsic parameters. 
Intrinsic parameters deal with the camera's internal characteristics, such as, its focal length, skew, distortion, and image center. 
Extrinsic parameters describe its position and orientation in the world. 
Knowing intrinsic parameters is an essential first step for 3D computer vision, as it allows you to estimate the scene's structure in Euclidean space and removes lens distortion, 
which degrades accuracy,  \cite{CameraCalibration1}. 

Recall the projection equations. The homogeneous coordinates of point $\mathbf{O}$ in 3D space can be transformed
to the camera plane, with the camera projection matrix $\mathbf{P}$, which includes both extrinsic
and extrinsic parameters. 

\begin{equation}
\mathbf{O}_{image} = \mathbf{P}\mathbf{O}_{world}
\end{equation}

Remember, the projected coordinates need to be converted to a homogeneous form
to get the $u,v$ pixel locations in pixel coordinates. 
We can  do this by dividing the image coordinates by the $z$-component. Finally, $u$ and $v$ can
then be multiplied with an arbitrary scale $s$. We multiply by $s$, as it will be useful
later on when we formulate the calibration problem. It is important to
note that scale plays a challenging role in understanding monocular
image information, as once points are projected from the 3D world onto
the 2D image plane, scale information is lost. Points in 3D space along
a ray from the camera center, all project to the same location
on the image plane, and it is therefore not possible to directly associate a depth to a point given only image information. 

\subsection{The Camera Calibration Problem}
\label{cam_calib_problem}

The camera calibration problem is defined as finding these unknown intrinsic and
extrinsic camera parameters, shown here in red given $n$ known 3D point coordinates and
their corresponding projection to the image plane. 

\begin{equation}
\begin{bmatrix}
su \\
sv \\
s
\end{bmatrix} =
\begin{bmatrix}
? & ? & ? & ? \\
? & ? & ? & ? \\
? & ? & ? & ?
\end{bmatrix}=
\begin{bmatrix}
X \\
Y \\
Z  \\
1
\end{bmatrix}
\end{equation}

Our approach will comprise of getting the $\mathbf{P}$ matrix first and
then decomposing it into the intrinsic parameters $\mathbf{K}$ and 
the extrinsic rotation parameters $\mathbf{R}$ and translation parameters $\mathbf{t}$. For calibration, we use a scene with
known geometry to get the location of our 3D points from the 2D image, resolving the scale
issue by measuring the actual 3D distance between the points that
are observed in the image. The most commonly used example would be a 3D checkerboard, with squares of
known size providing a map of fixed point locations to observe. We define our word
coordinate frame, in yellow and compute our 3D point coordinates and
their projections in the image. Associating 3D points to 2D projections can be
done either manually, by clicking on the purple points, for example or automatically, with checkerboard detectors. We can then set up
a system of equations to solve for the unknown parameters of $\mathbf{P}$. 

Now, let us form the system of linear equations that
needs to be solved. First, we expand
the projection equations to three equations through
matrix multiplication. 

\begin{eqnarray}
su = p_{11} X + p_{12} Y + p_{13} Z + p_{14} \\
sv = p_{21} X + p_{22} Y + p_{23} Z + p_{24} \\
s = p_{31} X + p_{32} Y + p_{33} Z + p_{34}
\end{eqnarray}

To get zero on the right-hand side of these equations, we move the right hand side to the left-hand side for each one. Then, we substitute
the third equation into equations one and two, and end up with
two equations per point. Therefore, if we have $n$ points, we have $2n$ associated equations. 
Putting these equations in matrix form gives us the following homogeneous
linear system. 

Since this is a homogeneous linear system, we can use the pseudo-inverse
or even better, the singular value decomposition to get the least
squares solution. Our simple linear
calibration approach has several advantages. It's easy to formulate, has a closed form solution, and often provides
really good initial points for non-linear calibration approaches. 

However, this simple approach has some disadvantages. One disadvantage
of solving for $\mathbf{P}$, is that we do not directly get the intrinsic and extrinsic
camera parameters. Furthermore, our linear model does not take into account
complex phenomena, such as radial and
tangential distortion. Finally, since we are solving via the linear least
squares method, we cannot impose constraints on our solution, such as requiring the focal
length to be non-negative. 

The camera projection matrix $\mathbf{P}$ by itself, is useful for projecting
3D points into 2D, but it has several drawbacks. It does not tell you
the cameras pose and it does not tell you about
the camera's internal geometry. Fortunately, we can
factorize $\mathbf{P}$ into intrinsic parameter matrix $\mathbf{K}$ 
and extrinsic rotation parameters $\mathbf{R}$ and translation parameters $\mathbf{t}$, 
using a linear algebra operation known as the $RQ$ factorization. Let us see how we perform
this factorization. 

\subsection{$RQ Factorization}
\label{rq_factorization}

First, we alter the representation of $\mathbf{P}$ to be a function of
the camera center $\mathbf{C}$. $\mathbf{C}$ is the point that projects to
zero when multiplied by $\mathbf{P}$. We multiply $\mathbf{K}$ into the matrix
to form two sub-matrices, $\mathbf{KR} and $-\mathbf{KRC}$. 
We will refer to the combination of $\mathbf{K}$ and $\mathbf{R}$ as the $\mathbf{M}$ matrix. 
We can now express our projection matrix $\mathbf{P}$ as 

\begin{equation}
\mathbf{P} = [\mathbf{M} | - \mathbf{MC}]
\end{equation}

From here, we use the fact that any square matrix can be factored into an upper
triangular matrix $\mathbf{R}$ and an orthogonal basis
to decompose $\mathbf{M}$ into upper triangular $\mathbf{R}$ and orthogonal basis $\mathbf{Q}$. 
In linear algebra, this procedure is known as $RQ$ factorization, which is a variant
of the more commonly referred to $QR$ factorization. In $QR$ factorization, we have the orthogonal $\mathbf{Q}$
first and then the upper triangular $\mathbf{R}$. 

\begin{framed}
\begin{remark}
Note here that the $\mathbf{R}$ and the output
of $RQ$ factorization, is a different variable than our rotation matrix $\mathbf{R}$. 
\end{remark}
\end{framed}


Let us now see how we can use the output of $RQ$ factorization of
the matrix $\mathbf{M}$ to retrieve $\mathbf{K}, $\mathbf{R}$, and $\mathbf{t}$ by aligning
these two expressions. The intrinsic calibration matrix
$\mathbf{K}$ is the output $\mathbf{R}$ of the $RQ$ factorization of $\mathbf{M}$. 
The rotation matrix $\mathbf{R}$ is the orthogonal basis $\mathbf{Q}$. 
Finally, we can extract the translation vector directly from K in the last column of the $\mathbf{P}$ matrix. 

$RQ$ factorization is a great tool to compute $\mathbf{K}, $\mathbf{R}$, and $\mathbf{t}$ from
the camera $\mathbf{P} matrix. However, some mathematical assumptions need to be performed to guarantee a unique solution
for these matrices. 

%We will explore
%these assumptions in further detail with this lesson's practice
%Jupiter notebook. 

Monocular camera calibration is a well-established tool that has excellent implementations
in C++, Python and MATLAB. OpenCV camera calibration see \url{CameraCalibration2} 

\section{Summary}

So, to summarize. In this section, we discussed a method to identify the camera projection
matrix $\mathbf{P}$. This method is known as camera calibration.
We saw  that the matrix $\mathbf{P}$ can be factored into the following components

\begin{itemize}
\item Camera intrinsic matrix $\mathbf{K}$ 
\itme Camera extrinsic parameters $\mathbf{R}$ and $\mathbf{t}, 
\end{itemize}
This can bed one  through $RQ$ factorization. 

\section{Questions}

\section{Assignements}